---
title: "plhdbR Vignette"
author: "Fernando A. Campos"
date: "`r Sys.Date()`"
output:
  md_document:
    variant: markdown_github
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# plhdbR

`plhdbR` is a collection of tools to work with data from the [Primate Life History Database](https://plhdb.org/) (PLHDB).

_There are no life history or fertility data included in this package._

To use the tools, you must have access to the PLHDB. The site is currently accessible for the working group members only.

>Karen B. Strier, Jeanne Altmann, Diane K. Brockman, Anne M. Bronikowski, Marina Cords, Linda M. Fedigan, Hilmar Lapp, Xianhua Liu, William F. Morris, Anne E. Pusey, Tara S. Stoinski and Susan C. Alberts (2010). The Primate Life History Database: a unique shared ecological data resource. _Methods in Ecology and Evolution_, 1(2), 199–211. doi: [10.1111/j.2041-210X.2010.00023.x](http://doi.org/10.1111/j.2041-210X.2010.00023.x)


`plhdbR` aims to facillitate the analysis of data in the PLHDB. It will have three main sets of utilities:

* Functions for reading and error-checking the life history and fertility data

* Functions for calculating vital rates

* Functions for loading and analyzing climate data

The available functions will be documented below as they become available.


## Preparation

To use this package, you first need to install `devtools` with:
```{r, eval = FALSE}
    install.packages("devtools")
```

Then, you can install the latest development version from github:

```{r, eval = FALSE}
  library(devtools)
  devtools::install_github("camposfa/plhdbR")
```

```{r, echo = FALSE}
  library(plhdbR)
```

This package makes heavy use of the data manipulation packages [stringr](http://cran.r-project.org/package=stringr), [lubridate](http://cran.r-project.org/package=lubridate), [tidyr](http://cran.r-project.org/package=tidyr), and [dplyr](http://cran.r-project.org/package=dplyr). If not already installed, `plhdbR` will install and load these packages automatically. It also provides a convenient wrapper to load them all in one fell swoop:

```{r, message = FALSE, eval = FALSE}
  load_plhdb_packages()
```


## Functions for working with life history and fertility data

The functions `read_bio_table` and `read_fert_table` read csv files of biography and fertility, respectively, created by the download buttons for these tables [https://plhdb.org](https://plhdb.org/). These functions strip away blank lines and header lines, parse any date/time columns, and return a well-ordered `dplyr::tbl_df`. To pull all the data from a given table, use search criteria like 'Study.ID != 10'. Note that the data are **not** extensively error-checked at this stage. If you try to feed these functions a normal csv file, bad things might happen.

### Biography data

```{r}
  # Assuming your file is called "biography_2015_03_17.csv"
  lh <- read_bio_table("biography_2015_03_17.csv")
  summary(lh)
```

### Fertility data

```{r}
  # Assuming your file is called "fertility_2015_03_17.csv"
  fert <- read_fert_table("fertility_2015_03_17.csv")
  summary(fert)
```


## Utility functions for climate data

`plhdbR` includes functionality for pulling the most recent data for several large-scale climate oscillation indices from various online data repositories. Supported climate indices include:

- Dipole Mode index ("dmi")
- Multivariate ENSO Index ("mei")
- Oceanic Niño Index ("oni")
- Southern Oscillation Index ("soi")
- Pacific Decadal Oscillation ("pdo")
- Atlantic Multidecadal Oscillation ("amo")
- North Atlantic Oscillation ("nao")

The function `load_climate_index` returns a named list of monthly climate index data. Each element of the list is an object of class `dplyr::tbl_df`, an extension of R's data.frame

```{r}
  indices <- load_climate_index(c("nao", "mei"))
  
  indices
```

Squash them all together with:

```{r}
  summary(bind_rows(indices))
```

