---
title: "plhdbR Vignette"
author: "Fernando A. Campos"
date: "`r Sys.Date()`"
output:
  md_document:
    variant: markdown_github
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

[Back to Respository](https://github.com/camposfa/plhdbR)

# Functions for working with the PLHDB data tables

## Prepare workspace

```{r, message = FALSE, warning = FALSE} 
  library(plhdbR)
  load_plhdb_packages()
```

## Utility functions for climate data

`plhdbR` includes functions for pulling the most recent data for several large-scale climate oscillation indices from various online data repositories. Supported climate indices include:

- Dipole Mode index ("dmi")
- Multivariate ENSO Index ("mei")
- Oceanic Niño Index ("oni")
- Southern Oscillation Index ("soi")
- Pacific Decadal Oscillation ("pdo")
- Atlantic Multidecadal Oscillation ("amo")
- North Atlantic Oscillation ("nao")

The function `load_climate_index` returns a named list of monthly climate index data. Each element of the list is an object of class `dplyr::tbl_df`.

```{r}
  indices <- load_climate_index(c("nao", "mei"))
  
  indices
```

Squash all the list elements together to a single `tbl_df` with `dplyr::bind_rows`:

```{r}
  indices_df <- dplyr::bind_rows(indices)
  summary(indices_df)
```

