---
title: "plhdbR Vignette"
author: "Fernando A. Campos"
date: "`r Sys.Date()`"
output:
  md_document:
    variant: markdown_github
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

[Back to Respository](https://github.com/camposfa/plhdbR)

# Functions for working with the PLHDB data tables

## Prepare workspace

```{r, message = FALSE, warning = FALSE} 
  library(plhdbR)
  load_plhdb_packages()
```

## Reading data

The functions `read_bio_table` and `read_fert_table` read csv files of biography and fertility data, respectively, created by the download buttons for these tables on the [PLHDB website](https://plhdb.org/). These functions strip away blank lines and header lines, parse any date/time columns, and return a well-ordered `dplyr::tbl_df`, an extension of R's `data.frame`. To pull all the data from a given table, use search criteria like `'Study.ID != 10'`. Note that the data are **not** extensively error-checked at this stage. If you try to feed these functions a normal csv file, bad things might happen.

### Biography data

```{r}
  # Assuming your file is called "biography_2015_03_17.csv"
  lh <- read_bio_table("biography_2015_03_17.csv")
  summary(lh)
```

### Fertility data

```{r}
  # Assuming your file is called "fertility_2015_03_17.csv"
  fert <- read_fert_table("fertility_2015_03_17.csv")
  summary(fert)
```

## Error-checking data

The functions `find_bio_errors` and `find_fert_errors` scan the loaded biography and fertility data, respectively, looking for errors of various kinds. Each function takes as an argument the relevant `dplyr::tbl_df` generated by the `read_..._table` functions listed above. `find_bio_errors` will find dates that are unrealistic as well as duplicate entries for the same (Study.Id, Animal.Id) combination. These are returned as named list elements `$error_dates` and `$error_duplicates`. `find_fert_errors` only scans the date/time fields for errors, since there are multiple fertility entries for some individuals. This is returned in a list with named element `$error_dates`.

```{r}
  # Check the biography data for errors
  bio_errors <- find_bio_errors(lh)
  bio_errors$error_dates
  bio_errors$error_duplicates
  
  # There is currently an error for animal "AKI" in the Karisoke data
  # Look at AKI's data
  lh %>% filter(Animal.Id == "AKI") %>% glimpse()

  # Fix error
  lh[lh$Animal.Id == "AKI", ]$Max.Birth.Date <- ymd("2007-11-26")

  # Check again
  find_bio_errors(lh)$error_dates

  # Check the fertility data for errors
  fert_errors <- find_fert_errors(fert)
  fert_errors$error_dates
  
  # Currently no date/time errors in the fertility data
```

The function `find_mom_errors` checks to see if all the animals listed in Mom.Id
in the biography table for a given study have a corresponding record in Animal.Id.

```{r}
  find_mom_errors(lh)
```

We can see from this that there are 40 such records that need to be corrected.
